From Freebase to Wikidata: The Great Migration

### 2016-01-29 paper.tex
Reworked Section 3.6.

We also used Google's Knowledge Graph~\cite{singhal2012} to add about 100,000 further mappings
by also matching a~topic and an item if they share a~Wikipedia link with the Knowledge Graph item.
There is some potential to add more data by looking at functional relations
like \textit{``father''} and \textit{``mother''},
but these relations only apply to relatively small sets of data
(like well known families), and the functional assumptions may run into some edge cases.
For example, the \texttt{/people/person/parents} property in Freebase also covers stepparents,
which may create cases where a~person has more than one male or female \texttt{/people/person/parents}.

@@

For the mapping between Freebase and Wikidata properties, we have chosen to apply a~manual
approach with the help of the Wikidata community.
Note that a~mapping of types is not needed, since Wikidata has no real notion of types
(or, to be more specific, the property \textit{``instance of''} (P31) is just another
property and connects to items without any special status).
With the help of the community, we have been able to quickly map around 360~properties
(note that Wikidata had a~total of about 1,800 properties at the time when this work
has been conducted in the Summer of 2015, and Freebase has around 37,700 properties).
For the mapping of properties where the domain are other topics or literals,
this mapping provides the Wikidata property to use.

For \texttt{/people/person/parents}, we have implemented a~special case
in order to be able to map them to the Wikidata properties
\textit{``father''} (P22) and \textit{``mother''} (P25)
based on the person's gender.

The mapping of CVTs (see \autoref{fig:cvt-obama}) is more complicated,
because it is not possible to have a~\mbox{1-to-1} relationship
between Freebase and Wikidata properties: the CVT is linked to the subject topic by one property
and has properties pointing to its component values, whereas the Wikidata statement
has a~main property value group that is qualified by other such groups.
In order to map a~CVT to a~statement, we have to know which of the CVT's properties
should be used as the main value of the statement, with the others being mapped to qualifiers.
In order to keep a~simple mapping that associates one Wikidata property to one Freebase property,
we map both the property that links the topic to the CVT
and the main property of the CVT to the same Wikidata property. In a~second step, we
then use merge these two mappings int a~single Wikidata claim, on which we add the
other properties from the CVT.

For example, if we map the CVT shown in \autoref{fig:cvt-obama},
we first use the Wikidata property \textit{``position held''} (P39) to map
\texttt{/politician/government\_positions\_held} and \linebreak
\texttt{/government\_position\_held/office\_position\_or\_title}. \linebreak
Second, for the qualifiers, we use \textit{``start time''} (P580)
for \texttt{/government\_position\_held/from} and \textit{``end time''} (P582)
for \texttt{/government\_position\_held/to}.
With the resulting \linebreak property mapping, the created Wikidata statement
looks like the one presented in \autoref{fig:statement-obama}.

@@

For CVTs that include sources, we map the property that links the CVT
to the source as a~Wikidata reference instead of a~qualifier.
The hardest part then is to actually create the full topic mapping
by considering CVTs and the complex datatypes of Wikidata.
For CVTs, when we have a~triple whose value is a~CVT, we just retrieve it,
map all its component triples (while ignoring ones that cannot be mapped)
and apply the above mechanism to get its main value.
For values, we use the datatype information stored with Wikidata properties
in order to cast values to the right type.
For quantities, we assume that the amounts are precise,
as there is no precision information in Freebase.
For globe coordinates, we have hardcoded the mapping of the \texttt{/location/geocode} CVT
into the globe-coordinate value type of Wikidata
and use the same precision guessing algorithm as the one used by the Wikidata user interface.
For dates, we just map the date and time values
(encoded using XSD types~\cite{campbell2003xmlschema})
to the Wikidata ``time'' datatype. However, we remove dates before 1920 that have
a~precision higher than year because we do not know
if they are relative to the Julian or the Gregorian calendar.
In such cases, we are not able to fill the ``calendar'' value of the Wikidata ``time'' datatype.

@@

We also cannot compare the number of facts and the number of statements directly,
even considering the lower number of topics.
Freebase encodes its knowledge far more redundantly than Wikidata.
To illustrate this, properties in Freebase often have a~\emph{reverse} property
that is used in order to be able to traverse the Freebase graph easily in both directions.
For example, the property \texttt{/people/person/place\_of\_birth} has the corresponding
reverse property \linebreak \texttt{/location/location/people\_born\_here}
that encodes exactly the same semantic information.
Such reverse properties sometimes exist in Wikidata
(like \textit{``children''} for \textit{``father''} and \textit{``mother''}),
but are far less common and also not automatically handled.

>>>>

We also used Google's Knowledge Graph~\cite{singhal2012} to add about 100,000 further mappings
by also matching a~topic and an item if they share a~Wikipedia link with the Knowledge Graph item.
There is some potential to add more data by looking at functional relations
like \texttt{father} and \texttt{mother},
but these relations only apply to relatively small sets of data
(like well known families), and the functional assumptions may run into some edge cases.
For example, the \texttt{/people/person/parents} property in Freebase also covers stepparents,
which may create cases where a~person has more than one male or female \texttt{/people/person/parents}.

@@

For the mapping between Freebase and Wikidata properties, we have chosen to apply a~manual
approach with the help of the Wikidata community.
Note that a~mapping of types is not needed, since Wikidata has no real notion of types
(or, to be more specific, the property \texttt{instance of} (P31) is just another
property and connects to items without any special status).
With the help of the community, we have been able to quickly map around 360~properties
(note that Wikidata had a~total of about 1,800 properties in the summer of 2015 when this work
was conducted, and Freebase had around 37,700 properties).
For the mapping of properties where the domain are other topics or literals,
this mapping provides the Wikidata property to use.

For \texttt{/people/person/parents}, we implemented a~special case
in order to map them to the Wikidata properties
\texttt{father} (P22) and \texttt{mother} (P25)
based on the person's gender.

\break The mapping of CVTs (see \autoref{fig:cvt-obama}) is more complicated,
because it is not possible to have a~\mbox{$1$-to-$1$} relationship
between Freebase and Wikidata properties: the CVT is linked to the subject topic by one property
and has properties pointing to its component values, whereas the Wikidata statement
has a~main property value group that is qualified by other such groups.
In order to map a~CVT to a~statement, we have to know which of the CVT's properties
should be used as the main value of the statement, with the others being mapped to qualifiers.
In order to keep a~simple mapping that associates one Wikidata property to one Freebase property,
we map both the property that links the topic to the CVT
and the main property of the CVT to the same Wikidata property. In a~second step, we
then merge these two mappings int a~single Wikidata claim, on which we add the
other properties from the CVT.

For example, if we map the CVT shown in \autoref{fig:cvt-obama},
we first use the Wikidata property \texttt{position held} (P39) to map
\texttt{/politician/government\_positions\_held} and \linebreak
\texttt{/government\_position\_held/office\_position\_or\_title}. \linebreak
Second, for the qualifiers, we use \texttt{start time} (P580)
for \texttt{/government\_position\_held/from} and \texttt{end time} (P582)
for \texttt{/government\_position\_held/to}.
With the then resulting \linebreak property mapping, the created Wikidata statement
looks like the one presented in \autoref{fig:statement-obama}.

@@

For CVTs that include sources, we map the property that links the CVT
to the source as a~Wikidata reference instead of a~qualifier.
The hardest part then is to actually create the full topic mapping
by considering CVTs and the complex datatypes of Wikidata.
For CVTs, when we have a~triple whose value is a~CVT, we just retrieve it,
map all its component triples (while ignoring ones that cannot be mapped),
and apply the above mechanism to get its main value.
For values, we use the datatype information stored with Wikidata properties
in order to cast values to the right type.
For quantities, we assume that the amounts are precise,
as there is no precision information in Freebase.
For globe coordinates, we have hardcoded the mapping of the \texttt{/location/geocode} CVT
into the globe-coordinate value type of Wikidata
and use the same precision guessing algorithm as the one used by the Wikidata user interface.
For dates, we just map the date and time values
(encoded using XSD types~\cite{campbell2003xmlschema})
to the Wikidata \texttt{time} datatype. However, we remove dates before 1920 that have
a~precision higher than year because we do not know
if they are relative to the Julian or the Gregorian calendar.
In such cases, we are unable to fill the \texttt{calendar} value of the Wikidata \texttt{time} datatype.

@@

We also cannot compare the number of facts and the number of statements directly,
even considering the lower number of topics.
Freebase encodes its knowledge far more redundantly than Wikidata.
To illustrate this, properties in Freebase often have a~\emph{reverse} property
that is used in order to be able to traverse the Freebase graph easily in both directions.
For example, the property \texttt{/people/person/place\_of\_birth} has the corresponding
reverse property \linebreak \texttt{/location/location/people\_born\_here}
that encodes exactly the same semantic information.
Such reverse properties sometimes exist in Wikidata
(like \texttt{children} for \texttt{father} and \texttt{mother}),
but are far less common and also not automatically handled.
